Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}

### 2016-07-20 Paper_zh.tex
翻译完成“6 交易执行”章节

\section{Transaction Execution} \label{ch:transactions}

The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$. It is assumed that any transactions executed first pass the initial tests of intrinsic validity. These include:

\begin{enumerate}
\item The transaction is well-formed RLP, with no additional trailing bytes;
\item the transaction signature is valid;
\item the transaction nonce is valid (equivalent to the sender account's current nonce);
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
\end{enumerate}

Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
\begin{equation}
\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
\end{equation}

Thus $\boldsymbol{\sigma}'$ is the post-transactional state. We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.

\subsection{Substate} 
Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction. We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
\begin{equation}
A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
\end{equation}

The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion. $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends). Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value. Though not immediately refunded, it is allowed to partially offset the total execution costs.

For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
\begin{equation}
A^0 \equiv (\varnothing, (), 0)
\end{equation}

\subsection{Execution}
We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
{} & + G_{transaction}
\end{align}

where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call. $G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}. $G$ is fully defined in Appendix \ref{app:fees}.

@@

The up-front cost $v_0$ is calculated as:
\begin{equation}
v_0 \equiv T_g T_p + T_v
\end{equation}

The validity is determined as:
\begin{equation}
\begin{array}[t]{rcl}
S(T) & \neq & \varnothing \quad \wedge \\
\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
g_0 & \leqslant & T_g \quad \wedge \\ 
v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
\end{array}
\end{equation}

Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.

The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$. The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$. The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.

We define the checkpoint state $\boldsymbol{\sigma}_0$:
\begin{eqnarray}
\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
\end{eqnarray}

Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
\begin{equation}
(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
\end{cases}
\end{equation}

where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_g - g_0
\end{equation}
and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.

Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.

After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
\begin{equation}
g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
\end{equation}

The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.

The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$. So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
\begin{eqnarray}
\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
m & \equiv & {B_H}_c
\end{eqnarray}

The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
\end{eqnarray}

And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
\begin{eqnarray}
\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
\end{eqnarray}

These are used to help define the transaction receipt, discussed later.

>>>>

\section{交易执行} \label{ch:transactions}

交易执行是以太坊协议最复杂的部分：它定义了状态转换函数 $\Upsilon$。任何交易的执行都需要先通过初始状态的内在合法性验证。这包括：

\begin{enumerate}
\item 交易以正确的RLP形式表示，没有多余字节跟在后面；
\item 交易的签名正确；
\item 交易的nonce正确（和发送账号当前nonce相同）；
\item gasLimit不能小于交易用到的gas数量$g_0$；
\item 发送者账户余额不小于需要预付的花费$v_0$。
\end{enumerate}

我们这样表达$\Upsilon$函数，其中$T$是交易，$\boldsymbol{\sigma}$是状态：
\begin{equation}
\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
\end{equation}

$\boldsymbol{\sigma}'$是交易执行后的状态。同时我们定义 $\Upsilon^g$ 为计算该交易执行所需gas的函数，$\Upsilon^\mathbf{l}$为计算该交易累计日志条目的函数。我们后续会给出正式定义。

\subsection{子状态} 
在交易执行过程中，我们会收集一些信息，这些信息在交易执行之后会被用到。我们称这些信息为\textit{子状态（transaction substate）}, 用$A$表示。子状态是个元组：
\begin{equation}
A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
\end{equation}

元组中包含一个自杀集$A_\mathbf{s}$，它是一组在交易完成后即会被丢弃的账户。$A_\mathbf{l}$是日志，是归档并索引过的虚拟机代码执行的`检查点（checkpoint）'，检查点允许以太坊之外的观察者（比如去中心化应用的前端）方便地追踪合约的调用情况。最后是存储退款额（refund balance）$A_r$，使用{\small SSTORE} 指令释放合约存储空间后退款额会变大。存储退款额虽然不是立即生效，但可以降低合约执行过程中的总花费。

为了简化，我们定义$A^0$为空子状态，它没有自杀集，没有日志，并且退款额为零：
\begin{equation}
A^0 \equiv (\varnothing, (), 0)
\end{equation}

\subsection{执行}
我们这样定义交易执行前需要预先支付的gas的数量$g_0$：
\begin{align}
g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{如} \quad i = 0 \\ G_{txdatanonzero} & \text{否则} \end{cases} \\
{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{否则} \end{cases} \\
{} & + G_{transaction}
\end{align}

这里 $T_\mathbf{i},T_\mathbf{d}$ 是交易的EVM的初始代码或数据，取决于该交易是合约生成类型的交易还是消息调用类型的交易。如果是合约生成类型的交易，$G_\text{txcreate}$会被累加上去；但如果交易不是合约生成型的，或者交易是在{\it 过度到家园（Homestead）分叉}之前执行的，$G_\text{txcreate}$则不会被累加上去。在附录\ref{app:fees}对$G$做了完整定义。

@@

需要预付的花费 $v_0$定义如下：
\begin{equation}
v_0 \equiv T_g T_p + T_v
\end{equation}

正确性的定义如下:
\begin{equation}
\begin{array}[t]{rcl}
S(T) & \neq & \varnothing \quad \wedge \\
\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
g_0 & \leqslant & T_g \quad \wedge \\ 
v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
\end{array}
\end{equation}

注意最后一个条件：该交易的油费上限$T_g$加上该区块之前所有油费消耗之和$\ell(B_\mathbf{R})_u$，不能大于该区块的 \textbf{gasLimit} ${B_H}_l$。

一个合法交易的执行伴随着对状态的不可撤销的变更：发送者账户$S(T)$的nonce被加一，账户余额被减去预付款额$T_gT_p$。可用于继续计算的油费$g$等于$T_g - g_0$。无论是合约生成还是消息调用，计算都会结束于一个最终状态（有可能和当前状态一样，这也是合法的）。状态更改是确定性的且保证是合法的：后续不会因此而存在非法交易。

我们定义检查点状态（checkpoint state） $\boldsymbol{\sigma}_0$为：
\begin{eqnarray}
\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{但是} \\
\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
\end{eqnarray}

我们用$\boldsymbol{\sigma}_P$表示交易后的临时状态。基于$\boldsymbol{\sigma}_0$去计算$\boldsymbol{\sigma}_P$取决于该交易是合约生成还是消息调用。我们定义交易后临时状态$\boldsymbol{\sigma}_P$，剩余油费$g'$和子状态 $A$构成的三元组为：
\begin{equation}
(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{如} \quad T_t = \varnothing \\
\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{否则}
\end{cases}
\end{equation}

其中$g$是扣除合约存在所需支付的基本花费后的剩余油费：
\begin{equation}
g \equiv T_g - g_0
\end{equation}

$T_o$表示原始交易人（original transactor）。如果消息调用或合约生成不是通过交易出发，而是通过EVM代码的执行而触发的，原始交易人和发送者（sender）就不相同。

注意，我们用$\Theta_{3}$表示方法值中只有前三个被使用。[不知道如何翻译：the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.]

消息调用或者合约生成处理完成后，会从剩余油费$g'$中用同样的价格计算退出款额$g^*$，并累加上退款计数器中的数值。
\begin{equation}
g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
\end{equation}

退款总额是剩余油费$g'$加上存储退款额$A_r$，但$A_r$被限制到不能超过实际油费消耗$T_g - g'$的一半。

购买油费所用的以太会被奖励给旷工，旷工由当前块$B$的受益人所指定。我们通过临时状态$\boldsymbol{\sigma}_P$定以终前状态（pre-final state）$\boldsymbol{\sigma}^*$：
\begin{eqnarray}
\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{但是} \\
\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
m & \equiv & {B_H}_c
\end{eqnarray}

删除自杀集中的所有账户后，就得到了最终状态$\boldsymbol{\sigma}'$：
\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{但是} \\
\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
\end{eqnarray}

最后我们定义该交易的耗油总量$\Upsilon^g$，以及该交易生成的日志$\Upsilon^\mathbf{l}$为：
\begin{eqnarray}
\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
\end{eqnarray}

它们可以帮助定义我们后续要讲的交易收据。
